### 2025-01-06 sections/1-introduction.tex
moved some floats for better whitespace usage

$\rho$-NN search also has a variety of applications.
For example, one could search for all genomes within a maximum edit distance of a query genome to find evolutionarily-related organisms~\cite{budowski2010fragbag}.
One could also search for all words within a maximum edit distance of a misspelled word to suggest corrections~\cite{ukkonen1985algorithms}.
Given an advertisement and a database of user profiles, one could search for all users whose profiles are ``similar enough'' to target the advertisement~\cite{zhang2020privacy}.
GPS and other location-based services use $\rho$-NN search to find nearby points of interest~\cite{zhang2020privacy} to a user's location.
%In all of these cases, approximate $\rho$-NN search may be sufficient, but exact $\rho$-NN search is preferable.

This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for exact $k$-NN search.
We also present improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
We provide a comparison to several current algorithms for similarity search; namely, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
We further benchmark CAKES on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, RadioML~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
Finally, we compare real-world datasets to an artificially-generated dataset obeying simple statistical properties.

>>>>

This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy-Scaling Search), a set of three novel algorithms for exact $k$-NN search.
We also compare CAKES against several current algorithms; namely, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
We further benchmark CAKES on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, RadioML~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.
Finally, we compare real-world datasets to an artificially-generated dataset obeying simple statistical properties.

### 2025-01-06 sections/2-methods.tex
starting to address major comments
moved some floats for better whitespace usage

The choice of an appropriate distance function varies by dataset and domain.
For example, with neural-network embeddings, one could use Euclidean (L2) or cosine distance.
With genomic or proteomic sequence data, Levenshtein, Smith-Waterman, Needleman-Wunsch or Hamming distances are useful.
With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.

CAKES derives its advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
While we sometimes use Euclidean notions to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
CAKES exploits the low LFD of such datasets to accelerate search.
We define LFD at some length scale around a point in the dataset as:

\begin{equation}
    \text{LFD}(q, r_1, r_2) = \frac{\text{log} \left( \frac{|B(q, r_1)|}{|B(q, r_2)|} \right) }{\text{log} \left( \frac{r_1}{r_2} \right) }
    \label{eq:methods:lfd-original}
\end{equation}

where $B(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.

We use a simplified version of this definition by calculating the LFD at a length scale where $r_1 = 2 \cdot r_2$.

\begin{equation}
    \text{LFD}(q, r) = \text{log}_2 \left( \frac{|B(q, r)|}{|B(q, \frac{r}{2})|} \right)
    \label{eq:methods:lfd-half}
\end{equation}

@@

\begin{algorithm} % enter the algorithm environment
    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[0] % enter the algorithmic environment
        \Require $C$, a cluster
        \Require $criteria$, user-specified continuation criteria

@@

% TODO: Move most of this to the supplement
% \subsection{\texorpdfstring{$\rho$}{p}-Nearest Neighbors Search}
% \label{sec:methods:rnn-search}

% We conduct $\rho$-NN search as described in~\cite{ishaq2019clustered}, but with the following improvement:
% when a cluster overlaps with the query ball, instead of always proceeding to search both of its children, we proceed only with those children that might contain points in the query ball.

% \begin{figure}[ht!]
%     \centering
%     \includegraphics[scale=0.75]{images/geometry/overlapping-children-3.pdf}
%     \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children. Here, $l$ is the left pole, $r$ is the right pole, and $q$ is the query. Other points and distances are described in the text.}
%     \label{fig:methods:overlapping-children}
%     \caption{CAKES uses geometric properties of clusters.}
% \end{figure}

% To determine whether both children can contain points in the query ball, we consider Figure~\ref{fig:methods:overlapping-children}.
% Here, we overload the notation for $\overline{x y}$ to refer both to the line segment joining points $x$ and $y$ as well as to the length of that line segment.

% Let $q$ denote the query, $\rho$ denote the search radius, and $l$ and $r$ denote the cluster's left and right poles respectively (see Section~\ref{sec:methods:clustering:building-the-tree}).
% Without loss of generality, we assume that $\overline{q r \vphantom{l}} \leq \overline{q l}$.
% Now let $q'$ be the projection of $q$ onto $\overline{l r}$, $m$ be the midpoint of $\overline{l r}$, and $d$ be the distance from $q'$ to $m$.
% As a consequence of how we assign a point in the parent cluster to the left child in Algorithm~\ref{alg:methods:partition}, if $\rho < d$, then the left child cannot contain points inside the query ball.
% In such a case we proceed to search only the right child.
% Otherwise, we proceed with both children.

% To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
% Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
% By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
% Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
% Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
% Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.

% Thus, $d \leq \rho \iff (\overline{q l} + \overline{q r \vphantom{l}})(\overline{q l} - \overline{q r \vphantom{l}}) \leq 2 \cdot \overline{l r} \cdot \rho$. Note, in particular, that this only requires distances between actual points from the dataset, and so it can be used with any distance function, even when $q'$ and $m$ are not actual points or cannot be imputed from the data.

% To perform $\rho$-NN search, we first perform a coarse \textit{tree-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:tree-search}, to find the leaf clusters that overlap with the query ball or any clusters which lie entirely within the query ball.
% Then, for all such clusters, we perform a finer-grained \textit{leaf-search}, as outlined in Algorithm~\ref{alg:methods:rnn-search:leaf-search}, to find all points that are no more than a distance $\rho$ from the query.
% The asymptotic complexity of $\rho$-NN is the same as in~\cite{ishaq2019clustered} and shown in Equation~\ref{eq:methods:rnn-search-complexity}.

% \begin{gather}
%     \mathcal{O}
%     \Bigg(
%         \underbrace{
%             \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
%         }_{\textrm{tree-search}}
%         \ + \
%         \underbrace{
%             \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
%             \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
%         }_{\textrm{leaf-search}}
%     \Bigg)
%     \label{eq:methods:rnn-search-complexity}
% \end{gather}
% where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B_X(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
    \caption{
        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
    }
    \label{fig:methods:deltas}
\end{figure}%

@@

This algorithm relies on the \textit{tree-search} and \textit{leaf-search} algorithms described in~\cite{ishaq2019clustered}.
We reproduce the algorithms here for completeness.

\begin{minipage}{.425\textwidth}
    \begin{algorithm}[H]\small
        \caption{tree-search($C$, $q$, $r$)}
        \label{alg:methods:rnn-search:tree-search}
        \begin{algorithmic}
            \Require $C$, a cluster
            \Require $q$, a query
            \Require $r$, a search radius

            \If{$C$ is a leaf \textbf{or} $\delta^+_C \leq r$}
                \State \textbf{return} $\{C\}$
            \Else
                \State $[l, r]$ $\Leftarrow$ \textit{pole}s of $C$
                \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
                \State \textbf{return} tree-search($L, q, r$)

                \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
            \EndIf

        \end{algorithmic}
    \end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
    \begin{algorithm}[H]\small
        \caption{leaf-search($Q$, $q$, $r$)}
        \label{alg:methods:rnn-search:leaf-search}
        \begin{algorithmic}
            \Require $Q$, a set of clusters
            \Require $q$, a query
            \Require $r$, a search radius

            \State $H \Leftarrow \emptyset$, a set of 2-tuples of points and their distances to the query

            \For{$C \in Q$}
                \If{$\delta^+_C \leq r$}
                    \State $H$ $\Leftarrow$ $H \cup \{C\}$
                \Else
                    \For{$p \in C$}
                        \If{$f(p, q) \leq r$}
                            \State $H$ $\Leftarrow$ $H \cup \{(p, f(p, q))\}$
                        \EndIf
                    \EndFor
                \EndIf
            \EndFor

            \State \textbf{return} $H$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}

For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) with a search radius $r$ equal to the radius of the root cluster divided by the cardinality of the dataset.
If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.

Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
In particular, we increase the radius by a factor of

\begin{equation}
    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)
    \label{eq:methods:repeated-rnn-factor}
\end{equation}

where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$.
We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.

@@

Once $\sum_{C \in Q} |C| \geq k$, we have found at least $k$ neighbors, and so we can stop increasing the radius.
We now perform \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) to find the $k$ nearest neighbors among the points in the clusters found during the previous steps.

\begin{algorithm}\small % enter the algorithm environment
    \caption{Repeated $\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
        \Require $root$, the root cluster
        \Require $q$, a query
        \Require $k$, the number of neighbors to find

@@

In ~\cite{ishaq2019clustered}, we showed that the complexity of $\rho$-NN search is

\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \
        \underbrace{
            \overbrace{ \big| B(q, \rho) \big|}^{\textrm{output size}}
            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
        }_{\textrm{leaf-search}}
    \Bigg)
    \label{eq:methods:rnn-search-complexity}
\end{gather}
where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.

@@

In addition, by our assumption that the LFD at the query is not significantly different from the LFD of nearby clusters, we have that $\hat{d} \approx d$.
By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has an asymptotic complexity of
\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \
        \underbrace{
            \overbrace{k}^{\textrm{output size}}
            \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
        }_{\textrm{leaf-search}}
    \Bigg)
    \label{eq:methods:repeated-rnn-complexity}
\end{gather}
where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
We note that the scaling factor should be close to 1 unless fractal dimension is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).

@@

The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is:

\begin{equation}
    \mathcal{O} \Big( ( \mathcal{T} + \mathcal{L} ) \log ( \mathcal{T} + \mathcal{L} ) \Big).
    \label{eq:methods:breadth-first-sieve-complexity}
\end{equation}

For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the priority queue operations.
Thus, the complexity is:

\begin{equation}
    \mathcal{O} \Big( \mathcal{T} \log \mathcal{T} + \mathcal{L} \log k \Big).
    \label{eq:methods:depth-first-sieve-complexity}
\end{equation}

>>>>

The choice of an appropriate distance function varies by dataset and domain.
For example, with neural-network embeddings, one could use Euclidean (L2) or cosine distance.
With genomic or proteomic sequence data, Levenshtein or Hamming distance is useful.
With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.

CAKES derives its advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
While we sometimes use Euclidean notions to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
CAKES exploits the low LFD of such datasets to accelerate search.
We define LFD at some length scale around a point in the dataset as:
\begin{equation}
    \text{LFD}(q, r_1, r_2) = \frac{\text{log} \left( \frac{|B(q, r_1)|}{|B(q, r_2)|} \right) }{\text{log} \left( \frac{r_1}{r_2} \right) }
    \label{eq:methods:lfd-original}
\end{equation}
where $B(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.

We use a simplified version of Equation~\ref{eq:methods:lfd-original} by using a length scale where $r_1 = 2 \cdot r_2$.
\begin{equation}
    \text{LFD}(q, r) = \text{log}_2 \left( \frac{|B(q, r)|}{|B(q, \frac{r}{2})|} \right)
    \label{eq:methods:lfd-half}
\end{equation}

@@

\begin{algorithm} % enter the algorithm environment
    \caption{Partition($C$, $criteria$)} % give the algorithm a caption
    \label{alg:methods:partition} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
        \Require $C$, a cluster
        \Require $criteria$, user-specified continuation criteria

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale=0.65]{images/geometry/deltas.pdf}
    \caption{
        {\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.
        ${\color{blue}\delta} = f(q, c)$ is the distance from the query to the cluster center $c$.
        ${\color{red}\delta^{+}} = \delta + r$ is the distance from the query to the theoretically farthest point in $C$.
        ${\color{green}\delta^{-}} = \text{max}(0, \delta - r)$ is the distance from the query to the theoretically closest point in $C$.
    }
    \label{fig:methods:deltas}
\end{figure}

\begin{minipage}{.425\textwidth}
    \begin{algorithm}[H]
        \caption{tree-search($C$, $q$, $r$)}
        \label{alg:methods:rnn-search:tree-search}
        \begin{algorithmic}
            \Require $C$, a cluster
            \Require $q$, a query
            \Require $r$, a search radius
            \If{$\delta^+_C \leq r$}
                \State \textbf{return} $\{C\}$
            \Else
                \State $[L, R]$ $\Leftarrow$ \textit{children} of $C$
                \State \textbf{return} tree-search($L, q, r$) \\
                \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ $\cup$ tree-search($R, q, r$)
            \EndIf
        \end{algorithmic}
    \end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
    \begin{algorithm}[H]
        \caption{leaf-search($Q$, $q$, $r$)}
        \label{alg:methods:rnn-search:leaf-search}
        \begin{algorithmic}
            \Require $Q$, a set of clusters
            \Require $q$, a query
            \Require $r$, a search radius
            \State $H \Leftarrow \emptyset$
            \For{$C \in Q$}
                \If{$\delta^+_C \leq r$}
                    \State $H$ $\Leftarrow$ $H \cup C$
                \Else
                    \For{$p \in C$}
                        \If{$f(p, q) \leq r$}
                            \State $H$ $\Leftarrow$ $H \cup \{p\}$
                        \EndIf
                    \EndFor
                \EndIf
            \EndFor
            \State \textbf{return} $H$
        \end{algorithmic}
    \end{algorithm}
\end{minipage}

@@

@@

This algorithm relies on the \textit{tree-search} (Algorithm~\ref{alg:methods:rnn-search:tree-search}) and \textit{leaf-search} (Algorithm~\ref{alg:methods:rnn-search:leaf-search}) as described in~\cite{ishaq2019clustered} and reproduced here for completeness.

For Repeated $\rho$-NN search, we begin by performing \textit{tree-search} with a search radius $r$ equal to the radius of the root cluster divided by the cardinality of the dataset.
If no clusters are found, then we double $r$ and perform tree-search again, repeating until we find at least one cluster.

Now, so long as $\sum_{C \in Q} |C| < k$, we continue to perform tree-search, but instead of doubling $r$ on each iteration, we multiply it by a factor determined by the LFD in the vicinity of the query ball.
In particular, we increase the radius by a factor of
\begin{equation}
    \min \left(2, \left( {\frac{k}{\sum_{C \in Q} |C|}} \right)^{\mu^{-1}} \right)
    \label{eq:methods:repeated-rnn-factor}
\end{equation}
where $\mu$ is the harmonic mean of the LFD of the clusters in $Q$.
We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.

@@

Once $\sum_{C \in Q} |C| \geq k$, we have found at least $k$ neighbors, and so we can stop increasing the radius.
We now perform \textit{leaf-search} to find the $k$ nearest neighbors among the points in the clusters found during the previous steps.

\begin{algorithm} % enter the algorithm environment
    \caption{Repeated $\rho$-NN($root$, $q$, $k$)} % give the algorithm a caption
    \label{alg:methods:repeated-rnn} % and a label for \ref{} commands later in the document
    \begin{algorithmic} % enter the algorithmic environment
        \Require $root$, the root cluster
        \Require $q$, a query
        \Require $k$, the number of neighbors to find

@@

In ~\cite{ishaq2019clustered}, we showed that the complexity of $\rho$-NN search is
\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \
        \underbrace{
            \overbrace{ \big| B(q, \rho) \big|}^{\textrm{output size}}
            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
        }_{\textrm{leaf-search}}
    \Bigg)
    \label{eq:methods:rnn-search-complexity}
\end{gather}
where $\hat{r}$ is the \textit{mean} radius of leaf clusters, $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy at that radius, $B(q, \rho)$ is a ball of radius $\rho$ around the query $q$, and $d$ is the LFD around the query at the length scale of $\rho$ and $\rho + 2 \cdot \hat{r}$.

@@

In addition, by our assumption that the LFD at the query is not significantly different from the LFD of nearby clusters, we have that $\hat{d} \approx d$.
By combining the bounds for tree-search and leaf-search, we see that Repeated $\rho$-NN has an asymptotic complexity of
\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \
        \underbrace{
            \overbrace{k}^{\textrm{output size}} \cdot
            \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
        }_{\textrm{leaf-search}}
    \Bigg)
    \label{eq:methods:repeated-rnn-complexity}
\end{gather}
where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
Note that the scaling factor should be close to 1 unless LFD is highly variable in the region around the query (i.e.,\,if $\hat{d}$ differs significantly from $d$).

@@

The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is
\begin{equation}
    \mathcal{O} \Big( ( \mathcal{T} + \mathcal{L} ) \log ( \mathcal{T} + \mathcal{L} ) \Big)
    \label{eq:methods:breadth-first-sieve-complexity}
\end{equation}

For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the priority queue operations.
Thus, the complexity is
\begin{equation}
    \mathcal{O} \Big( \mathcal{T} \log \mathcal{T} + \mathcal{L} \log k \Big)
    \label{eq:methods:depth-first-sieve-complexity}
\end{equation}
